---
title: "GettingCleaningDataProject"
author: "AP"
date: "4/25/2020"
output: html_document
---

The purpose of this project is to demonstrate your ability to collect, work with, and clean a data set.

One of the most exciting areas in all of data science right now is wearable computing - see for example this article . Companies like Fitbit, Nike, and Jawbone Up are racing to develop the most advanced algorithms to attract new users. The data linked to from the course website represent data collected from the accelerometers from the Samsung Galaxy S smartphone. A full description is available at the site where the data was obtained:
[http://archive.ics.uci.edu/ml/datasets/Human+Activity+Recognition+Using+Smartphones]{http://archive.ics.uci.edu/ml/datasets/Human+Activity+Recognition+Using+Smartphones}

Here are the data for the project:
[https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip]{https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip}

The project uses the following packeges .
```{r}
library('dplyr')
library('plyr')
```

Get file from the Web & and exctract it into a data frame using "read.table"
```{r}
url <- "https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip"
if(!file.exists("data.zip")){
    download.file(url, "data.zip")
}
unzip("data.zip")
```


Files in this projects:
* 'README.txt'
* 'features_info.txt': Shows information about the variables used on the feature vector.
* 'features.txt': List of all features.
* 'activity_labels.txt': Links the class labels with their activity name.
* 'train/X_train.txt': Training set.
* 'train/y_train.txt': Training labels.
* 'test/X_test.txt': Test set.
* 'test/y_test.txt': Test labels.



4. Appropriately labels the data set with descriptive variable names. Prepare dataFrame name row
```{r}
namesData <- read.table("UCI HAR Dataset/features.txt", header=FALSE ,sep="")
namesDataVec <- as.vector(make.unique(gsub("[()]|-|,","",namesData$V2)))
```


1. Merges the training and the test sets to create one data set.
First we the test set and teh training set samples.
We also attach the Appropriate labels to the sets
```{r}
trainingSet <- read.table("UCI HAR Dataset/train/X_train.txt", header=FALSE ,sep="")
colnames(trainingSet) <- namesDataVec

testSet <- read.table("UCI HAR Dataset/test/X_test.txt", header=FALSE ,sep="")
colnames(testSet) <- namesDataVec
```

We combine the two datasets:
```{r}
tranTestSet<-rbind(trainingSet,testSet)
```


2. Extracts only the measurements on the mean and standard deviation for each measurement.
```{r}
logicalVec <- (grepl("mean|std",namesDataVec))
tranTestMeanStdSet <- tranTestSet[logicalVec,]
```

3. Uses descriptive activity names to name the activities in the data set
```{r}
activityData    <- read.table("UCI HAR Dataset/activity_labels.txt", header=FALSE ,sep="")

trainYSet       <- read.table("UCI HAR Dataset/train/y_train.txt", header=FALSE ,sep="")
activity        <- sapply( trainYSet$V1, function(x) activityData$V2[x] )
activityDf      <- data.frame(matrix(unlist(activity), nrow=length(activity), byrow=T))
colnames(activityDf) <- "ActivityType"
trainingSet     <-cbind(trainingSet,activityDf)

testYSet        <- read.table("UCI HAR Dataset/test/y_test.txt", header=FALSE ,sep="")
activity        <- sapply( testYSet$V1, function(x) activityData$V2[x] )
activityDf      <- data.frame(matrix(unlist(activity), nrow=length(activity), byrow=T))
colnames(activityDf) <- "ActivityType"
testSet         <-cbind(testSet,activityDf)
```


5. From the data set in step 4, creates a second, independent tidy data set with the average of each variable for each activity and each subject.
```{r}
testSubjectSet  <- read.table("UCI HAR Dataset/test/subject_test.txt", header=FALSE ,sep="")
trainSubjectSet <- read.table("UCI HAR Dataset/train/subject_train.txt", header=FALSE ,sep="")
colnames(testSubjectSet) <- "Subject"
colnames(trainSubjectSet) <- "Subject"
testSet         <-cbind(testSet,testSubjectSet)
trainingSet     <-cbind(trainingSet,trainSubjectSet)
```

Finally, we calculte the mean of each variable in both sets
```{r}
testSetMeanGroup <- testSet     %>% group_by(ActivityType, Subject) %>%  summarise_if(is.numeric, mean, na.rm = TRUE)
str(testSetMeanGroup)
```
```{r}
trainingSetMeanGroup <- trainingSet %>% group_by(ActivityType, Subject) %>%  summarise_if(is.numeric, mean, na.rm = TRUE)
str(trainingSetMeanGroup)
```

### Code for Running rmarkdown from consolue: *rmarkdown::render("GettingCleaningDataProject.Rmd")*.